---
title: "CAUTI Study Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    navbar:
      - {title: "Overview", href: "#overview"}
      - {title: "Descriptives", href: "#descriptives"}
      - {title: "Survival", href: "#survival"}
      - {title: "Model", href: "#model"}
      - {title: "Data", href: "#data"}
runtime: shiny
---

```{r setup, message=FALSE, warning=FALSE}
# ----------------------------
# Packages
# ----------------------------
library(readxl)
library(tidyverse)
library(lubridate)

library(shiny)
library(flexdashboard)
library(DT)
library(plotly)

library(survival)
library(survminer)
library(broom)

library(gt)

# ----------------------------
# Data path (edit if needed)
# ----------------------------
DATA_PATH <- "E:/dataset/Cauti Study data 4-2024.xlsx"
# Portable option: put the xlsx in the same folder as this .Rmd and use:
# DATA_PATH <- "Cauti Study data 4-2024.xlsx"

# ----------------------------
# Build dataset ONCE (not reactive)
# ----------------------------
build_cauti4 <- function(path){
  stopifnot(file.exists(path))
  cauti <- read_excel(path)

  ID <- cauti %>% 
    distinct(PAT_MRN_ID) %>%
    mutate(ID = row_number())

  new_cauti <- cauti %>%
    left_join(ID, by = "PAT_MRN_ID") %>% 
    select(-"PAT_MRN_ID") %>% 
    select(ID, everything())

  cauti1 <- new_cauti %>% 
    filter(!PAT_AGE <= 17) %>% 
    filter(Gender != "Unknown", !is.na(ID)) %>% 
    select(-ORDER_FOR_FOLEY) %>% 
    filter(!(PLACEMENT_DATE <= HOSP_ADMSN_TIME - days(3))) %>% 
    mutate(
      REMOVAL_DATE = ifelse(
        REMOVAL_DATE > HOSP_DISCH_TIME | REMOVAL_DATE < PLACEMENT_DATE,
        HOSP_DISCH_TIME,
        REMOVAL_DATE
      ),
      PLACEMENT_DATE = ifelse(
        PLACEMENT_DATE < HOSP_ADMSN_TIME | PLACEMENT_DATE > HOSP_DISCH_TIME,
        HOSP_ADMSN_TIME,
        PLACEMENT_DATE
      )
    ) %>% 
    mutate(
      REMOVAL_DATE = as.POSIXct(REMOVAL_DATE, format = "%m/%d/%Y %H:%M"),
      PLACEMENT_DATE = as.POSIXct(PLACEMENT_DATE, format = "%m/%d/%Y %H:%M"),
      HOSP_ADMSN_TIME = as.POSIXct(HOSP_ADMSN_TIME, format = "%m/%d/%Y %H:%M"),
      HOSP_DISCH_TIME = as.POSIXct(HOSP_DISCH_TIME, format = "%m/%d/%Y %H:%M")
    )

  cauti2 <- cauti1 %>%
    mutate(
      Gender = if_else(Gender == "Male", 0, 1),
      DIABETES_ON_PROBLEM_LIST = if_else(DIABETES_ON_PROBLEM_LIST == "Y", 1, 0),
      RENAL_INSUFFICIENY_ON_PROBLEM_LIST = if_else(RENAL_INSUFFICIENY_ON_PROBLEM_LIST == "Y", 1, 0),
      Bowel_Incontinence = if_else(Bowel_Incontinence == "Y", 1, 0),
      Urinary_Incontinence = if_else(Urinary_Incontinence == "Y", 1, 0),
      CAUTI_LDA = if_else(CAUTI_LDA == "Y", 1, 0),
      CAUTI_Date = as.Date(CAUTI_Infection_date, format = "%m/%d/%Y"),
      admissionday = round(as.numeric(difftime(HOSP_DISCH_TIME, HOSP_ADMSN_TIME, units = "days")) + 1),
      CHG = round((CHG_NUMERATOR / CHG_DENOMINATOR) * 100),
      chg_100 = ifelse(CHG >= 100, 100, CHG),
      meatal = round((Meatal_Numerator / Meatal_Denominator) * 100),
      meatal_100 = ifelse(meatal >= 100, 100, meatal)
    ) %>%
    mutate(
      BMI_30 = ifelse(PAT_BMI < 30, 0, 1),
      BMI_cat = cut(PAT_BMI, breaks = c(-Inf, 18.5, 25, 30, Inf),
                    labels = c("underweight", "healthy weight", "overweight", "obese"))
    ) %>%
    mutate(
      Age_50 = ifelse(PAT_AGE < 50, 0, 1),
      Age_Cat = cut(PAT_AGE, breaks = c(-Inf, 50, 60, 70, 80, Inf),
                    labels = c("<50", "50-59", "60-69", "70-79", ">=80"))
    ) %>%
    mutate(
      CHG_Cat = cut(CHG, breaks = c(0, 49, 79, 100), labels = c("0-49", "50-79", "80-100")),
      Meatal_Cat = cut(meatal, breaks = c(0, 49, 79, 100), labels = c("0-49", "50-79", "80-100"))
    ) %>%
    filter(!(CAUTI_LDA == 1 & PLACEMENT_DATE > CAUTI_Date)) %>%
    select(-IP_LDA_ID, -Foley_Placement_unit, -LINE_DESCRIPTION , -LINE_TYPE) %>%
    mutate(CAUTI_LDA2 = ifelse(ConsecutiveFoley == "N", "0", CAUTI_LDA)) %>%
    mutate(foleyday = round(as.numeric(difftime(REMOVAL_DATE, PLACEMENT_DATE, units = "days")) + 1))

  combine_rows <- function(df) {
    combined <- data.frame()
    i <- 1
    while (i <= nrow(df)) {
      current_row <- df[i, ]
      if (i < nrow(df)) {
        next_row <- df[i + 1, ]
        if (next_row$PLACEMENT_DATE - current_row$REMOVAL_DATE <= days(1)) {
          combined_row <- current_row
          combined_row$REMOVAL_DATE <- next_row$REMOVAL_DATE
          combined <- rbind(combined, combined_row)
          i <- i + 2
        } else {
          combined <- rbind(combined, current_row)
          i <- i + 1
        }
      } else {
        combined <- rbind(combined, current_row)
        i <- i + 1
      }
    }
    combined
  }

  cleaned <- cauti2 %>%
    group_by(ID) %>%
    do(combine_rows(.))

  cauti3 <- cleaned %>% 
    mutate(r_date = lag(REMOVAL_DATE), .after = REMOVAL_DATE) %>% 
    arrange(ID, HOSP_ADMSN_TIME) %>%
    group_by(ID, HOSP_ADMSN_TIME) %>%
    mutate(
      num_foleys = cumsum(!duplicated(PLACEMENT_DATE)),
      foley_gap = ifelse(row_number() == 1, 0,
                         as.numeric(difftime(PLACEMENT_DATE, r_date, units = "days"))),
      consec_foley = cumsum(ifelse(row_number() == 1, 1,
                                   ifelse(foley_gap > 1, 1, 0))),
      foley_gap = round(foley_gap)
    ) %>% 
    mutate(consec_foley = ifelse(CAUTI_LDA == 1, 0, consec_foley)) %>% 
    mutate(duration_of_hospitalization = round(as.numeric(difftime(HOSP_DISCH_TIME, HOSP_ADMSN_TIME, units = "days")))) %>% 
    filter(!foleyday >= 50) %>% 
    ungroup()

  cauti4 <- cauti3 %>% 
    mutate(
      Bowel_Incontinence = as.numeric(Bowel_Incontinence),
      Urinary_Incontinence = as.numeric(Urinary_Incontinence)
    ) %>% 
    mutate(across(c(Bowel_Incontinence, Urinary_Incontinence), ~replace(., is.na(.), 0))) %>% 
    mutate(Gender = as.factor(Gender))

  cauti4
}

cauti4_df <- build_cauti4(DATA_PATH)

filtered_data <- reactive({
  d <- cauti4_df
  if (!is.null(input$gender) && input$gender != "All") d <- d %>% filter(Gender == input$gender)
  if (!is.null(input$agecat) && input$agecat != "All") d <- d %>% filter(Age_Cat == input$agecat)
  if (!is.null(input$cauti) && input$cauti != "All") d <- d %>% filter(CAUTI_LDA == as.integer(input$cauti))
  d
})

fmt_pct <- function(x) sprintf("%.2f%%", x)
``` 

Sidebar {.sidebar}
-------------------------------------

```{r}
selectInput("gender", "Gender", choices = c("All", "0", "1"), selected = "All")
selectInput("agecat", "Age Category", choices = c("All", levels(cauti4_df$Age_Cat)), selected = "All")
selectInput("cauti", "CAUTI", choices = c("All", "0", "1"), selected = "All")
helpText("Gender coding: 0 = Male, 1 = Female")
```

# Overview {#overview}

Row
-------------------------------------

### Key Metrics

```{r}
renderValueBox({ valueBox(n_distinct(filtered_data()$ID), caption = "Unique Patients") })
renderValueBox({ valueBox(nrow(filtered_data()), caption = "Foley Episodes") })
renderValueBox({ valueBox(fmt_pct(mean(filtered_data()$CAUTI_LDA, na.rm = TRUE) * 100), caption = "CAUTI Rate") })
renderValueBox({ valueBox(round(median(filtered_data()$foleyday, na.rm = TRUE), 1), caption = "Median Foley Days") })
```

Row {data-height=560}
-------------------------------------

### Filtered Cohort Overview

```{r}
render_gt({
  d <- filtered_data()
  tab <- d %>%
    summarize(
      `Unique Patients` = n_distinct(ID),
      `Mean Age` = round(mean(PAT_AGE, na.rm = TRUE), 1),
      `Mean BMI` = round(mean(PAT_BMI, na.rm = TRUE), 1),
      `Male (n)` = sum(Gender == "0", na.rm = TRUE),
      `Female (n)` = sum(Gender == "1", na.rm = TRUE),
      `CAUTI Yes (n)` = sum(CAUTI_LDA == 1, na.rm = TRUE),
      `CAUTI No (n)` = sum(CAUTI_LDA == 0, na.rm = TRUE)
    ) %>%
    pivot_longer(everything(), names_to = "Metric", values_to = "Value")

  gt(tab) %>% tab_header(title = "Filtered Cohort Overview")
})
```

# Descriptives {#descriptives}

Row {data-height=620}
-------------------------------------

### Foley Days Distribution

```{r}
renderPlotly({
  d <- filtered_data()
  p <- ggplot(d, aes(x = foleyday)) +
    geom_histogram(bins = 30) +
    labs(x = "Foley Days", y = "Count", title = "Distribution of Foley Days") +
    theme_minimal()
  ggplotly(p)
})
```

### Age Distribution by CAUTI

```{r}
renderPlotly({
  d <- filtered_data() %>% mutate(CAUTI = factor(CAUTI_LDA, levels = c(0,1), labels = c("No", "Yes")))
  p <- ggplot(d, aes(x = CAUTI, y = PAT_AGE, fill = CAUTI)) +
    geom_boxplot() +
    labs(x = "CAUTI", y = "Age", title = "Age Distribution by CAUTI Status") +
    theme_minimal()
  ggplotly(p)
})
```

Row {data-height=620}
-------------------------------------

### CAUTI by CHG Category

```{r}
renderPlotly({
  d <- filtered_data() %>% mutate(CAUTI = factor(CAUTI_LDA, levels = c(0,1), labels = c("No", "Yes")))
  p <- d %>%
    count(CHG_Cat, CAUTI) %>%
    ggplot(aes(x = CHG_Cat, y = n, fill = CAUTI)) +
    geom_col(position = "dodge") +
    labs(x = "CHG Category", y = "Count", title = "CAUTI by CHG Category") +
    theme_minimal()
  ggplotly(p)
})
```

### CAUTI by Meatal Care Category

```{r}
renderPlotly({
  d <- filtered_data() %>% mutate(CAUTI = factor(CAUTI_LDA, levels = c(0,1), labels = c("No", "Yes")))
  p <- d %>%
    count(Meatal_Cat, CAUTI) %>%
    ggplot(aes(x = Meatal_Cat, y = n, fill = CAUTI)) +
    geom_col(position = "dodge") +
    labs(x = "Meatal Category", y = "Count", title = "CAUTI by Meatal Care Category") +
    theme_minimal()
  ggplotly(p)
})
```

# Survival {#survival}

Row {data-height=760}
-------------------------------------

### Kaplan–Meier Curve (Overall)

```{r}
renderPlot({
  d <- filtered_data()
  km_fit <- survfit(Surv(foleyday, CAUTI_LDA) ~ 1, data = d)
  ggsurvplot(
    km_fit,
    data = d,
    conf.int = TRUE,
    risk.table = TRUE,
    ggtheme = theme_minimal(),
    title = "Kaplan–Meier Curve (Foley Days)",
    xlab = "Days",
    ylab = "Survival Probability"
  )
})
```

Row {data-height=760}
-------------------------------------

### Kaplan–Meier Curve by Gender

```{r}
renderPlot({
  d <- filtered_data()
  validate(need(length(unique(d$Gender)) > 1, "Select All genders (or include both 0 and 1) to compare KM curves."))
  km_fit_gender <- survfit(Surv(foleyday, CAUTI_LDA) ~ Gender, data = d)
  ggsurvplot(
    km_fit_gender,
    data = d,
    pval = TRUE,
    conf.int = TRUE,
    risk.table = TRUE,
    ggtheme = theme_minimal(),
    title = "Kaplan–Meier Curves by Gender (0=Male, 1=Female)",
    xlab = "Days",
    ylab = "Survival Probability",
    legend.labs = c("Male", "Female")
  )
})
```

# Model {#model}

Row {data-height=700}
-------------------------------------

### Logistic Regression (Adjusted Odds Ratios)

```{r}
render_gt({
  d <- filtered_data()

  model <- glm(
    CAUTI_LDA ~ PAT_AGE + Gender + PAT_BMI + DIABETES_ON_PROBLEM_LIST + 
      RENAL_INSUFFICIENY_ON_PROBLEM_LIST + Bowel_Incontinence + 
      Urinary_Incontinence + Num_Foley_Urine_Culture_Collected + 
      duration_of_hospitalization + num_foleys + foleyday + meatal_100 + chg_100,
    data = d,
    family = binomial
  )

  or_tab <- broom::tidy(model, conf.int = TRUE, exponentiate = TRUE) %>%
    filter(term != "(Intercept)") %>%
    mutate(
      Predictor = case_when(
        term == "PAT_AGE" ~ "Age (years)",
        term == "Gender1" ~ "Gender (Female vs Male)",
        term == "PAT_BMI" ~ "BMI",
        term == "DIABETES_ON_PROBLEM_LIST" ~ "Diabetes",
        term == "RENAL_INSUFFICIENY_ON_PROBLEM_LIST" ~ "Renal insufficiency",
        term == "Bowel_Incontinence" ~ "Bowel incontinence",
        term == "Urinary_Incontinence" ~ "Urinary incontinence",
        term == "Num_Foley_Urine_Culture_Collected" ~ "# Foley urine cultures",
        term == "duration_of_hospitalization" ~ "Hospital days",
        term == "num_foleys" ~ "# Foley episodes",
        term == "foleyday" ~ "Foley days",
        term == "meatal_100" ~ "Meatal care (%)",
        term == "chg_100" ~ "CHG (%)",
        TRUE ~ term
      ),
      OR = estimate,
      `CI 95%` = sprintf("%.3f to %.3f", conf.low, conf.high),
      `p-value` = signif(p.value, 3)
    ) %>%
    select(Predictor, OR, `CI 95%`, `p-value`)

  gt(or_tab) %>% tab_header(title = "Adjusted Odds Ratios (Logistic Regression)")
})
```

# Data {#data}

Row {data-height=820}
-------------------------------------

### Data Preview (first 500 rows)

```{r}
renderDT({
  d <- filtered_data() %>%
    select(ID, PAT_AGE, Gender, PAT_BMI, DIABETES_ON_PROBLEM_LIST,
           RENAL_INSUFFICIENY_ON_PROBLEM_LIST, Bowel_Incontinence,
           Urinary_Incontinence, CHG_Cat, Meatal_Cat, foleyday,
           duration_of_hospitalization, num_foleys, CAUTI_LDA) %>%
    head(500)

  datatable(d, options = list(pageLength = 15, scrollX = TRUE))
})
